---
title: "sensitivity analysis on models"
author: "Shuyu, Marame, Kevin, Théo, Florence, Florent"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=FALSE, results="hide")
source("common.R")
if (! exists("mrun_stage_all_models")) {mrun_stage_all_models = memoise::memoise(run_stage_all_models)}
# apply_func = function(...){epimedtools::monitored_apply(mod=1, ...)}
if (! exists("apply_func")) {apply_func = apply}

```

# Simualtion params 

We generate simulation parameters from `r kc` dataset.

We compare obtained indicators between `r kc` and simulates dataset.

In both cases we see that HR associated with msm13 transition is not stable.

```{r data kc, echo=TRUE, results="verbatim"}
if (!exists("kc")) {kc = "BRCA"}
# data = data[data$type%in%kc,]
data = openxlsx::read.xlsx("TCGA-CDR-SupplementalTableS1.xlsx", sheet = "TCGA-CDR", na.strings="#N/A")
d = dfnew_creation(data)
# table(d$type, d$stage)
d = d[d$type%in%kc,]
d$dss = survival::Surv(d$DSS.time, d$DSS)
d$pfi = survival::Surv(d$PFI.time, d$PFI)

p = generate_params_for_sim(d, kc=kc)
p = p_orig = signif(p,2)
print(p)
dfsimu = dfsimu_creation_3times(p)
dsimu = dfnew_creation(dfsimu)
dsimu$stage = dfsimu[rownames(dsimu),]$stage
table(dsimu$stage)
dsimu$dss = survival::Surv(dsimu$DSS.time, dsimu$DSS)
dsimu$pfi = survival::Surv(dsimu$PFI.time, dsimu$PFI)

results      = mrun_stage_all_models(d)
results_simu = mrun_stage_all_models(dsimu)

layout(matrix(1:2,1), respect=TRUE)
results$type=kc
plot_hrs(results, main=results$type[1])
results_simu$type=paste0(kc, "_simu")
plot_hrs(results_simu, main=results_simu$type[1])

p = p_orig = signif(p,1)
```

# Screen 1: `censor_param`

We see that the HRs are globally stretched with `censor_param`

But the HR associated with msm13 transition stays independently unstable.


```{r screen1, results="verbatim", echo=TRUE}
layout(matrix(1:6,2), respect=TRUE)

for (censor_param in seq(0.02, 0.22, by=0.04)) {
  p[9] = censor_param

  dfsimu = dfsimu_creation_3times(p)
  d = dfnew_creation(dfsimu)
  d$stage = dfsimu[rownames(d),]$stage
  table(d$stage)
  d$dss = survival::Surv(d$DSS.time, d$DSS)
  d$pfi = survival::Surv(d$PFI.time, d$PFI)

  results = mrun_stage_all_models(d)

  print("T01")
  print(table(d$T01))
  print("T02")
  print(table(d$T02))
  print("T12")
  print(table(d$T12))
  plot_hrs(results, main=paste0("censor_param: ", censor_param))  
}
```

# Screen 2: size of each group

We see that the accuracy of the HR is not symmetrically affected by the proportion of the stage.

```{r screen2, results="verbatim", echo=TRUE}
layout(matrix(1:6,2), respect=TRUE)
for (prop in seq(0.1, 0.9, by=0.16)) {
  p[9] = 0.06
  p[1] = 1000 * prop
  p[5] = 1000 * (1-prop)

  dfsimu = dfsimu_creation_3times(p)
  d = dfnew_creation(dfsimu)
  d$stage = dfsimu[rownames(d),]$stage
  table(d$stage)
  d$dss = survival::Surv(d$DSS.time, d$DSS)
  d$pfi = survival::Surv(d$PFI.time, d$PFI)

  results = mrun_stage_all_models(d)

  plot_hrs(results, main=paste0(p[1],"/", p[5]))  
}
```

# Screen 3: transitions


```{r screen3, results="verbatim", echo=TRUE}
p = p_orig 
p[2:4]
p[6:8]
layout(matrix(1:6,2), respect=TRUE)

i02s = c(.01,.03,.05)

p = p_orig 
params = expand.grid(
  nlo          = 500 ,
  i01lo        = p[2],
  i02lo        = i02s,
  i12lo        = p[4],
  nhi          = 500 ,
  i01hi        = p[6],
  i02hi        = i02s,
  i12hi        = p[8],
  censor_param = 0.06,
  seed = 1:3
)


stats = apply_func(params, 1, function(p){
  # p = params[4,]
  dfsimu = dfsimu_creation_3times(p=unlist((p[1:9])), seed=p[[10]])
  d = dfnew_creation(dfsimu)
  d$stage = dfsimu[rownames(d),]$stage
  table(d$stage)
  results = mrun_stage_all_models(d)
  # plot_hrs(results, main=paste0(p, collapse=" "))
  ret = c(p, msm13=results["msm13",3] - results["msm13",2])
})
stats = data.frame(t(stats))
stats[!is.finite(stats$msm13),]$msm13 = max(stats[is.finite(stats$msm13),]$msm13)

layout(matrix(1:2,1), respect=TRUE)
boxplot(msm13~i02lo+i02hi, stats, las=2)
max_msm13 = 20
stats[stats$msm13>max_msm13,]$msm13 = max_msm13
boxplot(msm13~i02lo+i02hi, stats, las=2)

i02los        = c(.03, .05)
i02his        = c(.03, .05)

```




```{r screen4, results="verbatim", echo=TRUE}
p = p_orig 
p[2:4]
p[6:8]
i01s = c(.02,.03, .04, .06)
params = expand.grid(
  nlo          = 500 ,
  i01lo        = i01s,
  i02lo        = .03,
  i12lo        = p[4],
  nhi          = 500 ,
  i01hi        = i01s,
  i02hi        = .05,
  i12hi        = p[8],
  censor_param = 0.06,
  seed = 1:3
)

stats = apply_func(params, 1, function(p){
  # p = params[4,]
  # p = params[1,]
  dfsimu = dfsimu_creation_3times(p=unlist((p[1:9])), seed=p[[10]])
  d = dfnew_creation(dfsimu)
  d$stage = dfsimu[rownames(d),]$stage
  table(d$stage)
  results = mrun_stage_all_models(d)
  # plot_hrs(results, main=paste0(p, collapse=" "))s
  ret = c(p, msm13=results["msm13",3] - results["msm13",2])
})
stats = data.frame(t(stats))
if (any(!is.finite(stats$msm13))) {stats[!is.finite(stats$msm13),]$msm13 = max(stats[is.finite(stats$msm13),]$msm13)}

layout(matrix(1:2,1), respect=TRUE)
boxplot(msm13~i01lo+i01hi, stats, las=2)
max_msm13 = 20
stats[stats$msm13>max_msm13,]$msm13 = max_msm13
boxplot(msm13~i01lo+i01hi, stats, las=2)

i01los        = c(.02, .025)
i01his        = c(.02,.04,.06)
```


```{r screen5, results="verbatim", echo=TRUE}
p = p_orig 
p[2:4]
p[6:8]
i12s = c(.2,.3,.4)
params = expand.grid(
  nlo          = 500 ,
  i01lo        = .02,
  i02lo        = .03,
  i12lo        = i12s,
  nhi          = 500 ,
  i01hi        = .06,
  i02hi        = .05,
  i12hi        = i12s,
  censor_param = 0.06,
  seed = 1:3
)

stats = apply_func(params, 1, function(p){
  # p = params[4,]
  # p = params[1,]
  dfsimu = dfsimu_creation_3times(p=unlist((p[1:9])), seed=p[[10]])
  d = dfnew_creation(dfsimu)
  d$stage = dfsimu[rownames(d),]$stage
  table(d$stage)
  results = mrun_stage_all_models(d)
  # plot_hrs(results, main=paste0(p, collapse=" "))s
  ret = c(p, msm13=results["msm13",3] - results["msm13",2])
})
stats = data.frame(t(stats))
if (any(!is.finite(stats$msm13))) {stats[!is.finite(stats$msm13),]$msm13 = max(stats[is.finite(stats$msm13),]$msm13)}

layout(matrix(1:2,1), respect=TRUE)
boxplot(msm13~i12lo+i12hi, stats, las=2)
max_msm13 = 20
if (stats$msm13>max_msm13) {
  stats[stats$msm13>max_msm13,]$msm13 = max_msm13
  boxplot(msm13~i12lo+i12hi, stats, las=2)
}

i12los        = c(.2,.3)
i12his        = c(.2,.3)
```



```{r screen6, results="verbatim", echo=TRUE}
p = p_orig 
p[2:4]
p[6:8]
params = expand.grid(
  nlo          = 500 ,
  i01lo        = i01los,
  i02lo        = i02los,
  i12lo        = i12los,
  nhi          = 500 ,
  i01hi        = i01his,
  i02hi        = i02his,
  i12hi        = i12his,
  censor_param = 0.06,
  seed = 1:3
)



i01los
i02los
i12los
i01his
i02his
i12his


stats = apply_func(params, 1, function(p){
  # p = params[4,]
  # p = params[1,]
  dfsimu = dfsimu_creation_3times(p=unlist((p[1:9])), seed=p[[10]])
  d = dfnew_creation(dfsimu)
  d$stage = dfsimu[rownames(d),]$stage
  table(d$stage)
  results = mrun_stage_all_models(d)
  # plot_hrs(results, main=paste0(p, collapse=" "))s
  hrs = results[,1]
  dhrs = results[,3] - results[,2]
  names(hrs) = paste0("hr_", rownames(results))
  names(dhrs) = paste0("dhr_", rownames(results))
  hrs 
  ret = c(p, hrs, dhrs)
  ret
})
stats = data.frame(t(stats))
max_score = 20

for (key_score in colnames(stats)[19:26]) {
  f = formula(paste0(key_score, "~i01lo+i01hi+i02lo+i02hi+i12lo+i12hi"))  
  layout(matrix(1:2,1), respect=TRUE)
  if (any(!is.finite(stats[[key_score]]))) {stats[!is.finite(stats[[key_score]]),key_score] = max(stats[is.finite(stats[[key_score]]),key_score])}
  boxplot(f, stats, las=2, main=key_score)
  if (any(stats[[key_score]]>max_score)) stats[stats[[key_score]]>max_score,key_score] = max_score
  boxplot(f, stats, las=2, main=key_score)
}
```

# Correlations

```{r pairs, fig.height=9}
pairs(stats[,19:26])
pairs(stats[,11:18])
```

# PCA


```{r pca, eval=TRUE, message=FALSE, warning=FALSE, results="verbatim"}
# head(stats)
# mat = stats[,-c(1, 2, 5, 9, 10, 11:18, 19:26)]
mat0 = stats[,c(2, 3, 4, 6:8, 22)]
mat1 = stats[,c(2, 3, 4, 6:8, 11, 12)]
mat2 = stats[,c(2, 3, 4, 6:8, 13, 14, 15)]
mat3 = stats[,c(2, 3, 4, 6:8, 16, 17, 18)]
for (mat in list(mat0, mat1, mat2, mat3)) {
  head(mat)
  pca = prcomp(mat, scale=TRUE)
  v = pca$sdev * pca$sdev
  p = v / sum(v) * 100

  layout(matrix(1:6,2), respect=TRUE)
  # layout(matrix(1:2,1), respect=TRUE)
  barplot(p)
  for (i in 1:5) {
    j = i+1
    plot(pca$x[,i], pca$x[,j], xlab=paste0("PC", i, "(", signif(p[i], 3), "%)"), ylab=paste0("PC", j, "(", signif(p[j], 3), "%)"), pch=16)  
    scale_factor = min(abs(c(min(c(pca$x[,i], pca$x[,j])), max(c(pca$x[,i], pca$x[,j])))))  
    # scale_factor = min(abs(c(max(min(pca$x[,i]), min(pca$x[,j])), min(max(pca$x[,i]), max(pca$x[,j])))))
    arrows(0,0,pca$rotation[,i]*scale_factor, pca$rotation[,j]*scale_factor, col="grey")
    text(pca$rotation[,i]*scale_factor, pca$rotation[,j]*scale_factor, rownames(pca$rotation), col=2)
  }  
}
```

```{r}
i01los
i02los
i12los
i01his
i02his
i12his

p = c(
  nlo          = 500 ,
  i01lo        = min(i01los),
  i02lo        = max(i02los),
  i12lo        = max(i12los),
  nhi          = 500,
  i01hi        = max(i01his),
  i02hi        = min(i02his),
  i12hi        = min(i12his),
  censor_param = 0.06
)


p = c(
  nlo          = 500 ,
  i01lo        = min(i01los),
  i02lo        = min(i02los),
  i12lo        = min(i12los),
  nhi          = 500,
  i01hi        = max(i01his),
  i02hi        = min(i02his),
  i12hi        = min(i12his),
  censor_param = 0.06
)


p = c(
  nlo          = 500 ,
  i01lo        = min(i01los),
  i02lo        = min(i02los),
  i12lo        = min(i12los),
  nhi          = 500,
  i01hi        = min(i01his),
  i02hi        = max(i02his),
  i12hi        = min(i12his),
  censor_param = 0.06
)


dfsimu = dfsimu_creation_3times(p)
dsimu = dfnew_creation(dfsimu)
dsimu$stage = dfsimu[rownames(dsimu),]$stage
table(dsimu$stage)
dsimu$dss = survival::Surv(dsimu$DSS.time, dsimu$DSS)
dsimu$pfi = survival::Surv(dsimu$PFI.time, dsimu$PFI)

results = mrun_stage_all_models(dsimu)

layout(matrix(1:2,1), respect=TRUE)
plot_hrs(results, main=paste0(p, collapse=" "))



```


# Session Information

```{r results="verbatim"}
sessionInfo()
date()
```


















